---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Read activity.csv file from working directory
```{r}
activity <- read.csv("./activity.csv")
```
Load required packages
```{r}
library(lubridate)
library(ggplot2)
```
Change format of date variable
```{r}
activity$date <- ymd(as.character(activity$date))
```

## What is mean total number of steps taken per day?

Total number of steps taken per day:
```{r}
aggregate(steps ~ date, data=activity, sum)
```

Histogram of the total number of steps taken each day:
```{r histogram1}
stepsbyday <- aggregate(steps ~ date, data=activity, sum)
qplot(steps, data=stepsbyday, xlab="total steps per day", ylab="number of days")
```

Mean of the total number of steps taken per day:
```{r}
mean(stepsbyday$steps)
```

Median of the total number of steps taken per day:
```{r}
median(stepsbyday$steps)
```

## What is the average daily activity pattern?

Time series plot of the 5-minute interval, and the average number of steps taken, averaged across all days:
```{r timeseries1}
timeseries <- aggregate(steps ~ interval, data=activity, mean)
g <- ggplot(timeseries, aes(interval, steps))
g + geom_line() + ylab("average steps taken")
```

Which 5-minute interval contains the maximum number of steps?
```{r}
timeseries[which.max(timeseries$steps),]
```

## Imputing missing values

Total number of missing values in the dataset:
```{r}
sum(is.na(activity))
```

Create a new dataset (activity_imp) that is equal to the original dataset but with missing values for steps replaced by the mean value for that 5-minute interval:
```{r}
activity_imp <- activity
activity_imp$meansteps_int <- with(activity_imp, ave(steps, interval, FUN=function(x) mean(x, na.rm=TRUE)))
activity_imp$steps[is.na(activity_imp$steps)] = activity_imp$meansteps_int
```

Histogram of the total number of steps taken per day, after imputation:
```{r histogram2}
stepsbyday_imp <- aggregate(steps ~ date, data=activity_imp, sum)
qplot(steps, data=stepsbyday_imp, xlab="total steps per day", ylab="number of days")
```

Mean of the total number of steps taken per day, after imputation:
```{r}
mean(stepsbyday_imp$steps)
```
There is no impact of imputation on the mean number of steps per day

Median of the total number of steps taken per day, after imputation:
```{r}
median(stepsbyday_imp$steps)
```
There is a slight increase in median number of steps per day after imputation

## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset (weekday) indicating whether the given date is a weekday or weekend:
```{r}
activity_imp$weekday <- as.factor(ifelse(
                        weekdays(activity_imp$date) %in% c("Saturday", "Sunday"), "Weekend", "Weekday"))
```

Panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days:
```{r timeseries2}
timeseries_imp <- aggregate(steps ~ interval + weekday, data=activity_imp, mean)
g <- ggplot(timeseries_imp, aes(interval, steps))
g + geom_line() + facet_grid(weekday ~ .) + ylab("average steps taken")
```
